metric analysis

### 2021-05-29 metric_analysis.Rmd
Explanatory text

@@

@@

resist_combined %>% 
  pivot_longer(dlwrf:vgrd) %>% 
  ggplot(aes(trough_resistance, value, color = variable)) +
  geom_point() + 
  facet_wrap(variable~name, scales="free_y") + 
  geom_smooth(method = "lm", formula = y ~ x) +
  theme(legend.position = "none", axis.text = element_text(size = 5))

# Fit a Random Forest for each variable
rc_all <- split(resist_combined, resist_combined$variable)
rf_all <- lapply(rc_all, function(x) 
  randomForest(trough_resistance ~ ., data = x[-1:-2], importance = TRUE)
)
imp_all <- lapply(rf_all, function(x) {
  imp <- as.data.frame(importance(x))
  imp$met_var <- rownames(imp)
  as_tibble(imp)
})
importance <- bind_rows(imp_all, .id = "variable")
```

>>>>

Take a look at the distribution of met data, and perform a Principal Components
Analysis (PCA) to examine auto-correlation between the various met variables.

@@

# Met-metric data

We see from the met analysis above that vbdsf and vddsf are highly
correlated with nbdsf and nddsf respectively; remove one pair.

Take a first look at the relationship between output variables and met variables.

resist_combined %>% 
  pivot_longer(dlwrf:vgrd) %>% 
  ggplot(aes(trough_resistance, value, color = variable)) +
  geom_point() + 
  facet_wrap(variable~name, scales = "free_y") + 
  geom_smooth(method = "lm", formula = y ~ x) +
  theme(legend.position = "none", axis.text = element_text(size = 5))
```

@@

Fit a Random Forest model to each output variable (GPP, etc.) and then calculate
variable importance for each model. Make partial dependence plots for the
three most important met variables for each output variable.

```{r, message=FALSE}
# Fit a Random Forest for each variable (across all treatment severities)
# Note the r.f. functions don't play well with tibbles, so change to data frame
library(randomForest)
rc_all <- split(as.data.frame(resist_combined), resist_combined$variable)
rf_all <- lapply(rc_all, function(x) 
  randomForest(trough_resistance ~ ., data = x[-1:-2], importance = TRUE)
)

# Importance metrics and partial plot data
import <- list()
partial <- list()
for(var in names(rf_all)) {
  mod <- rf_all[[var]]
  imp <- as.data.frame(importance(mod))
  imp$met_var <- rownames(imp)
  import[[var]] <- as_tibble(imp)

  impvar <- rownames(imp)[order(imp[, 1], decreasing = TRUE)]
  for (i in seq_along(impvar[1:3])) {
    pp <- partialPlot(mod, rc_all[[var]], impvar[i], plot = FALSE)
    partial[[paste(var, i)]] <- tibble(var = var, 
                                       met_var = impvar[i],
                                       x = pp$x, y = pp$y)
  }
}

import <- bind_rows(import, .id = "var")
library(forcats)
lapply(split(import, import$var), function(x) {
  ggplot(x, aes(fct_reorder(met_var, `%IncMSE`), `%IncMSE`)) + 
    geom_col() +
    coord_flip() +
    xlab("Met variable") +
    ggtitle(unique(x$var))
})

partial <- bind_rows(partial)
ggplot(bind_rows(partial), aes(x, y, color = var)) + 
  geom_line() + 
  xlab("Met var value") + ylab("Metric") +
  ggtitle("Partial dependence plot") +
  facet_grid(var ~ met_var, scales = "free")
```

# Summary

Interestingly, `tmp` (air temperature) and `vbdsf` (sunlight) are among the
top-three most important drivers for all four output variables 
(GPP, NEP, NPP, Rh) for resistance. Atmospheric pressure (`pres`) is
among the top three for both GPP and NPP resistance: the former because of
stomatal effects perhaps (?), and the latter because
of NPP's strong dependence on GPP.
